New lower bounds for Schur and weak Schur numbers

### 2021-06-06 article/introduction.tex
CC des changements dans rapport vers article + modifs intro

\qquad We are interested in partioning the set of integers \(\{1, ..., p\}\) into \(n\) subsets such that there is no 
subset containing three integers \(x\), \(y\) and \(z\) verifying \(x + y = z\). We say these subsets are 
\textit{sum-free}. If we add the hypothesis \(x \neq y\), we say the subsets are \textit{weakly sum-free}. The 
greatest \(p\) for which there exists a partition into \(n\) sum-free subsets is called the \(n^{\text{th}}\) Schur 
number and is denoted \(S(n)\) \cite{Schur1917}. Likewise for weakly sum-free partitions we define \(\WS(n)\) 
the \(n^{\text{th}}\) weak Schur number \cite{Irving1973}. Only the first values of these sequences are known. 
This article focuses on the lower bounds for these numbers.

@@

\qquad Before Rowley's "template"-based approach for Schur and Ramsey numbers \cite{RowleyRamsey}, the 
previous generic construction for Schur numbers was given by Abbott and Hanson \cite{AbbottHanson} in 1972 
with a recursive construction. It used to give the best lower bounds for all sufficiently large numbers. No equivalent 
was known for weak Schur numbers and as a result the best known partitions for large weak Schur numbers 
did not utilize the weakly sum-free hypothesis.

\par
As for smaller weak Schur numbers, the best lower bounds were obtained by conducting a computer search. Eliahou 
\cite{EliahouBook}, Bouzy \cite{Bouzy2015AnAP} and Rafilipojaona \cite{Rafilipojaona} improved the lower 
bounds with Monte-Carlo methods. This was the main approach during the past decade. This search for weakly 
sum-free partitions relied on the recursive assumption that a good weakly sum-free partition into \(n+1\) colors 
starts with a good weakly sum-free partition into \(n\) colors. 
However, we show in the \hyperref[SearchSpace]{last section} that this search space may not contain the optimal 
solutions. Thus further works using similar methods might need to put that assumption aside and find a different 
search space.

\par
In 2020, Rowley introduced the notion of template for Schur and Ramsey numbers \cite{RowleyRamsey} which 
generalizes Abbott and Hanson's construction and gives new lower bounds (and inequalties) for Schur numbers. 
Rowley also gives two inequalities for weak Schur numbers \cite{RowleyWS} that yield significant improvements 
over previous lower bounds which besides do utilize the \textit{weakly} sum-free hypothesis.

@@

\begin{center}

\textbf{Table 1 - Comparison of lower bounds for Schur numbers}

\begin{tabular}{|*{13}{c|}}
    \hline
    \(n\) & 1 & 2 & 3 & 4 & 5 & 6 & 7 & 8 & 9 & 10 & 11 & 12 \\
    \hline
    \multirow{2}{*}{Before Rowley} & 1* & 4* & 13* & 44* & 160* & 536 & 1\,680 & 5\,041 & 15\,124 & 51\,120 & 172\,216 & 575\,664 \\
      & & & & & \cite{Heule2017} & \cite{Fredricksen} & \cite{Fredricksen} & \cite{AbbottHanson} & 
    \cite{AbbottHanson} & \cite{AbbottHanson} & \cite{AbbottHanson} & \cite{AbbottHanson} \\
    \hline
    \begin{tabular}{@{}c@{}}Rowley \\ \cite{RowleyRamsey}\end{tabular} & & & & & & & & 5\,286 & 17\,694 & 
    60\,320 & 201\,696 & 631\,840 \\
    \hline
    \hyperref[Schur]{\textbf{Our results}}  & & & & & & & & & \textbf{17\,803} & \textbf{60\,948} & \textbf{203\,828} & \textbf{638\,548} \\
    \hline
\end{tabular}

\vspace{2.5ex}

\textbf{Table 2 - Comparison of lower bounds for weak Schur numbers}

\begin{tabular}{|*{13}{c|}}
    \hline
    \(n\) & 1 & 2 & 3 & 4 & 5 & 6 & 7 & 8 & 9 & 10 & 11 & 12 \\
    \hline
    \multirow{2}{*}{Before Rowley} & 2* & 8* & 23* & 66* & 196 & 582 & 1\,740 & 5\,201 & 15\,596 & 51\,520 & 172\,216 & 575\,664 \\
      & & & & & \cite{ELIAHOU2012175} &\cite{EliahouBook} & \cite{Rafilipojaona} & \cite{Rafilipojaona} & 
    \cite{Rafilipojaona} & \cite{AbbottHanson} & \cite{AbbottHanson} & \cite{AbbottHanson} \\
    \hline
    \begin{tabular}{@{}c@{}}Rowley \\ \cite{RowleyWS}\end{tabular} & & & & & & 642 & 2\,146 & 6\,976 & 21\,848 
    & 70\,778 & 241\,282 & 806\,786 \\
    \hline
    \hyperref[WeakSchur]{\textbf{Our results}} & & & & & & & & & \textbf{22\,536} & \textbf{71\,214} & \textbf{243\,794} & \textbf{815\,314} \\
    \hline
\end{tabular}

\vspace{1.5ex}
* denotes an exact value, not just a lower bound
\end{center}

@@

\qquad The main contribution of this article is a generalization of the concept of template to weak Schur numbers. 
It gives new lower bounds (and inequalities) for weak Schur numbers. This construction also includes as a special 
case an analogous for weak Schur numbers of Abbott and Hanson's construction for Schur numbers.

\par
In the \hyperref[Schur]{first section}, we explain Rowley's template-based construction in the context of 
Schur numbers and then give new templates, thus providing new lowers bounds and inequalities as well as 
showing that the growth rates for both Schur and Ramsey numbers \(R_n(3)\) exceed 3.28.

\par
Then, in the \hyperref[WeakSchur]{second section}, we generalize the concept of templates to weak Schur numbers 
and provide new lower bounds for weak Schur numbers.

\par
Finally, in the \hyperref[SearchSpace]{third section}, we analyze the significant difference between new lower 
bounds obtained with templates and the former lower bounds obtained by computer search and we provide 
evidence which indicate that the main assumption made in those articles removes the optimal partitions from the 
search space.

>>>>

\qquad We are interested in partioning the set of integers \(\{1, ..., p\}\) into \(n\) subsets such that there is no 
subset containing three integers \(x\), \(y\) and \(z\) verifying \(x + y = z\). We say these subsets are 
\textit{sum-free}. If we add the hypothesis \(x \neq y\), we say the subsets are \textit{weakly sum-free}. The 
greatest \(p\) for which there exists a partition into \(n\) sum-free subsets is called the \(n^{\text{th}}\) Schur 
number and is denoted \(S(n)\) \cite{Schur1917}. Likewise for weakly sum-free partitions we define \(\WS(n)\) 
the \(n^{\text{th}}\) weak Schur number \cite{Irving1973}. Only the first values of these sequences are known. 
In this study, we improve the lower bounds on these numbers.

@@

\qquad Before Rowley's "template"-based approach for Schur and Ramsey numbers \cite{RowleyRamsey}, the 
previous generic construction for Schur numbers was given by Abbott and Hanson \cite{AbbottHanson} in 1972 
with a recursive construction. It used to give the best lower bounds for all sufficiently large numbers. No equivalent 
was known for weak Schur numbers and as a result the best known partitions for large weak Schur numbers 
did not use the weakly sum-free hypothesis.

\par
As for smaller weak Schur numbers, the best lower bounds were obtained by conducting a computer search. Eliahou 
\cite{EliahouBook}, Bouzy \cite{Bouzy2015AnAP} and Rafilipojaona \cite{Rafilipojaona} improved the lower 
bounds with Monte-Carlo methods. This was the main approach during the past decade. This search for weakly 
sum-free partitions relied on the recursive assumption that a good weakly sum-free partition into \(n+1\) subsets 
starts with a good weakly sum-free partition into \(n\) subsets.

\par
In 2020, Rowley introduced the notion of templates for Schur and Ramsey numbers \cite{RowleyRamsey} which 
generalizes Abbott and Hanson's construction and produces new lower bounds (and inequalties) for Schur numbers. 
He also provides two inequalities for weak Schur numbers \cite{RowleyWS} that yield significant improvements 
over previous lower bounds, and besides do utilize the \textit{weakly} sum-free hypothesis.

@@

\begin{table}
  \centering
\begin{tabular}{|*{13}{c|}}
    \hline
    \(n\) & 1 & 2 & 3 & 4 & 5 & 6 & 7 & 8 & 9 & 10 & 11 & 12 \\
    \hline
    \multirow{2}{*}{Before Rowley} & 1* & 4* & 13* & 44* & 160* & 536 & 1\,680 & 5\,041 & 15\,124 & 51\,120 & 172\,216 & 575\,664 \\
      & & & & & \cite{Heule2017} & \cite{Fredricksen} & \cite{Fredricksen} & \cite{AbbottHanson} & 
    \cite{AbbottHanson} & \cite{AbbottHanson} & \cite{AbbottHanson} & \cite{AbbottHanson} \\
    \hline
    \begin{tabular}{@{}c@{}}Rowley \\ \cite{RowleyRamsey}\end{tabular} & & & & & & & & 5\,286 & 17\,694 & 
    60\,320 & 201\,696 & 631\,840 \\
    \hline
    \hyperref[Schur]{\textbf{Our results}}  & & & & & & & & & \textbf{17\,803} & \textbf{60\,948} & \textbf{203\,828} & \textbf{638\,548} \\
    \hline
\end{tabular}
\caption{Comparison of lower bounds for Schur numbers}
\label{table:1}
\end{table}
\vspace{2.5ex}

\begin{table}
  \centering
\begin{tabular}{|*{13}{c|}}
    \hline
    \(n\) & 1 & 2 & 3 & 4 & 5 & 6 & 7 & 8 & 9 & 10 & 11 & 12 \\
    \hline
    \multirow{2}{*}{Before Rowley} & 2* & 8* & 23* & 66* & 196 & 582 & 1\,740 & 5\,201 & 15\,596 & 51\,520 & 172\,216 & 575\,664 \\
      & & & & & \cite{ELIAHOU2012175} &\cite{EliahouBook} & \cite{Rafilipojaona} & \cite{Rafilipojaona} & 
    \cite{Rafilipojaona} & \cite{AbbottHanson} & \cite{AbbottHanson} & \cite{AbbottHanson} \\
    \hline
    \begin{tabular}{@{}c@{}}Rowley \\ \cite{RowleyWS}\end{tabular} & & & & & & 642 & 2\,146 & 6\,976 & 21\,848 
    & 70\,778 & 241\,282 & 806\,786 \\
    \hline
    \hyperref[WeakSchur]{\textbf{Our results}} & & & & & & & & & \textbf{22\,536} & \textbf{71\,214} & \textbf{243\,794} & \textbf{815\,314} \\
    \hline
\end{tabular}

\vspace{1.5ex}
* denotes the exact value
\caption{Comparison of lower bounds for weak Schur numbers}
\label{table:2}
\end{table}

@@

\qquad Our main contribution is a generalization of the concept of template to weak Schur numbers. 
Our templates provide new lower bounds (and inequalities) for weak Schur numbers. This construction also includes, 
as a special case, a construction similar to Abbott and Hanson's \cite{AbbottHanson}, but this time with 
\textit{weak} Schur numbers.

\par
In the \hyperref[Schur]{section 2}, we explain Rowley's template-based construction in the context of 
Schur numbers. Then, we give new templates, thus providing new lowers bounds and inequalities as well as 
showing that the growth rates for both Schur and Ramsey numbers \(R_n(3)\) exceed 3.28.

\par
Then, in the \hyperref[WeakSchur]{section 3}, we generalize the concept of templates to weak Schur numbers 
and provide new lower bounds for weak Schur numbers.
